\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
clean sheet

\abstract{ We make progress in understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators at one-loop directly from Feynman rules. Consideration of
  this theory highlights that the lifting from one-loop
  color-kinematics dual numerators to two loops is just as challenging
  as the progression from tree-level to one-loop.  As a first step in
  understanding the two-loop challenge, we construct color-dual
  two-loop four-point nonlinear sigma model (NLSM) numerators using a
  $D$-dimensional color-dual bootstrap.  The double-copy of the
  constructed NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known
  physical unitarity cuts of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These
  NLSM numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our loop-level bootstrap, we also find that the most general \emph{local} ansatz
  for two-loop four-point pure Yang-Mills fails to satisfy
  global color-kinematics duality. We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
%\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
 % numerator

\item \ace{I am starting to get a bit concerned about the overuse of the
  phrase ``color-dual''} \nhp{We all choose capital $N$}
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological four-point two-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

@@

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation \be % [[19]]
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} for discussions of BELs in the context of
color-kinematics duality.

\ace{Awkward phrasing/transition.} \nhp{will message language} What has just been described is
sometimes referred to as ``global'' color-kinematics duality because
the Jacobi identities are solved by a single (global) integrand with
off-shell internal lines.  In order to perform the double copy it is
enough to simply satisfy the Jacobi identities on the generalized
unitarity cuts, meaning that each graph can be assigned its own ansatz
\cite{Bern:2015ooa}.  If a global solution cannot be found, this is
one way to enlarge the ansatz where the other option described above
was to incorporate non-local terms in the ``numerators''.  A globally
color-dual integrand is desirable because of its simplicity and
because a color-dual Lagrangian would naively produce precisely this
object.

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}. Recently, generalized unitarity was optimized for
even-multiplicity theories like NLSM \cite{Carrasco:2023qgz}.

Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{n[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.  \jm{Again, I love the colors!
  Would it be possible to add $\ell$ labels to the RHS of
  \cref{eq:example-cut-sum} to show what $\ell_5$, $\ell_6$... are?}
\ace{It is doable, will work on it once our text is a bit more happy}

@@

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of YZ-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply:
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}
\ee
As one can see, the kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

As noted above, we can select out can manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, the
$D$-dimensional description of $J$-theory is just the following blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed YZ theory matches to NLSM for off-shell
Y-particles. We provide explicit expressions for these one-loop amplitudes in the next section.

>>>>

\abstract{ We make progress in understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators at one-loop directly from Feynman rules. Consideration of
  this theory highlights that the lifting from one-loop
  color-kinematics dual numerators to two loops is just as challenging
  as the progression from tree-level to one-loop.  As a first step in
  understanding the two-loop challenge, we construct color-dual
  two-loop four-point nonlinear sigma model (NLSM) numerators using a
  $D$-dimensional color-dual bootstrap.  The double-copy of the
  constructed NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known
  physical unitarity cuts of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These
  NLSM numerators are in tension with generalizations our one-loop construction. Using our loop-level bootstrap, we also find that the most general \emph{local} ansatz
  for two-loop four-point pure Yang-Mills fails to satisfy
  global color-kinematics duality. We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

%TODO:\\
%For now, let's put the things we need to check/compute here in the abstract. 
%\begin{enumerate}
%\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
 % numerator

%\item \ace{I am starting to get a bit concerned about the overuse of the
  %phrase ``color-dual''} \nhp{We all choose capital $N$}
%\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
   % , \textcolor{cut-color}{\texttt{cut-color}},
  %  \textcolor{refl-color}{\texttt{refl-color}}, and
  %  \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
  %  other colors so that we don't have to keep changing colors in
   % multiple locations when we want to change diagram color schemes.
    %I think I have updated all of the diagrams to use the
  %  correctly-named color, but haven't changed any of the backing
   % colors for the names.}
%\end{enumerate}
}

@@

@@

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation \be % [[19]]
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} for discussions of BELs in the context of
color-kinematics duality.
\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological four-point two-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}
\ace{Awkward phrasing/transition.} \nhp{will message language} What has just been described is
sometimes referred to as ``global'' color-kinematics duality because
the Jacobi identities are solved by a single (global) integrand with
off-shell internal lines.  In order to perform the double copy it is
enough to simply satisfy the Jacobi identities on the generalized
unitarity cuts, meaning that each graph can be assigned its own ansatz
\cite{Bern:2015ooa}.  If a global solution cannot be found, this is
one way to enlarge the ansatz where the other option described above
was to incorporate non-local terms in the ``numerators''.  A globally
color-dual integrand is desirable because of its simplicity and
because a color-dual Lagrangian would naively produce precisely this
object.

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}, which was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{n[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.  \jm{Again, I love the colors!
  Would it be possible to add $\ell$ labels to the RHS of
  \cref{eq:example-cut-sum} to show what $\ell_5$, $\ell_6$... are?}
\ace{It is doable, will work on it once our text is a bit more happy}

@@

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply:
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}
\ee
As one can see, the kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

As noted above, we can select out can manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, the
$D$-dimensional description of $J$-theory is just the following blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
Y-particles. We provide explicit expressions for these one-loop amplitudes in the next section.

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.
